Moral Foundations of U.S. Political News Organizations

### 2019-04-07 pilot_manuscript.Rmd
Built the source descriptive tables and wrote about it in Exp1

####build the table here####
tableprint = matrix(NA, nrow=4, ncol=7)
colnames(tableprint) = c("Sources", "Unique_Words", "stuff")

tableprint[ , 1] = levels(final$Source)
tableprint[ , 2] = tapply(final$uniquewords, final$Source, mean)

#change this to match
apa_table(as.data.frame(tableprint),
          note = "For mean and standard deviation values, 'C' and 'L' refer to 'conservative' and 'liberal,' respectively",
          caption = "Experiment 1 Results - Multilevel Model", 
          col.names = c("Foundation", "$M_C$", "$SD_C$", "$M_L$", "$SD_L$", "$t$", "$p$", "$d$"),
          format = "latex",
          escape = F,
          align = c("l", rep("c", 6))
          )

overall valence for each source

types/tokens ratios, etc. for each source also number of articles

table - columns are types (unique words), tokens (total words), readability statistics, number of articles, total number of words, valence

talk about the table here

@@

The multilevel model did not indicate the presence of any significant or practical effect of political lean for any of the five moral foundations. The strongest effect size was observed for the *authority/respect* foundation, but the effect was in the opposite direction from what was originally hypothesized - liberal sources tended to use more *authority/respect* words than did conservative sources. Descriptive and test statistics, *p*-values and effect sizes (Cohen's *d*) can be found in Table 1. To interpret the weighted scores, one can examine the mean and standard deviations for each. A zero score for the mean, with a non-zero standard deviation, would indicate a perfect balance of positive and negative words in each category, likely representing a neutral tone when all words are considered. Negative percentages would indicate more representation of the negative words in the MFD area, while positive percentages indicate an endorsement of the positive words in a MFD. Therefore, we suggest using the sign of the mean score to determine the directionality of the endorsement for the MFD (positive, neutral, negative), and the standard deviation to ensure that a zero score is not zero endorsement (i.e., a SD of zero indicates no words were used). Based on the weighted percent values for the five foundations, the researchers observed that MFD words seem to make up a small portion of the article text. Furthermore, the observed percentages and means appear to indicate a generally positive endorsement of all five foundations across both liberal and conservative sources.

@@

@@

tapply(final_gs$FK, final_gs$Source, mean)
tapply(final_gs$FK, final_gs$Source, sd)
```

>>>>

####build the table here####
tableprint = matrix(NA, nrow=4, ncol=11)
colnames(tableprint) = c("Source", "Mean_Valence", "SD_Valence", "Number_of_Articles", "Total_Words", "Mean_Tokens", "SD_Tokens", "Mean_Types", "SD_Types", "Mean_Readability", "SD_Readability")

tableprint[ , 1] = print(unique(final$Source))
tableprint[ , 2] = tapply(final$valenceavg, final$Source, mean)
tableprint[ , 3] = tapply(final$valenceavg, final$Source, sd)
tableprint[ , 4] = tapply(final$X, final$Source, length)
tableprint[ , 5] = tapply(final$wordcount, final$Source, sum)
tableprint[ , 6] = tapply(final$wordcount, final$Source, mean)
tableprint[ , 7] = tapply(final$wordcount, final$Source, sd)
tableprint[ , 8] = tapply(final$uniquewords, final$Source, mean)
tableprint[ , 9] = tapply(final$uniquewords, final$Source, sd)
tableprint[ , 10] = tapply(final$FK, final$Source, mean)
tableprint[ , 11] = tapply(final$FK, final$Source, sd)

#change this to match
apa_table(as.data.frame(tableprint),
          note = "Readability statistics were calculated using the Flesch-Kincaid Grade Level readability formula.",
          caption = "Experiment 1 - Descriptive Statistics by Source", 
          col.names = c("Source", "$M_Valence", "$SD_Valence$", "Number of Articles", "Total Words", "$M_Tokens$", "$SD_Tokens$", "$M_Types$", "SD_Types", "$M_Readability$", "$SD_Readability$"),
          format = "latex",
          escape = F,
          align = c("l", rep("c", 6))
          )

The researchers calculated descriptive statistics for each news source in order to understand any and all fundamental linguistic differences in the sources' use of English. Statistics calculated included *z*-scored valence, number of articles per source, total number of words per source, average number of tokens (words) per article in each source, average number of types (unique words) per article in each source, and mean readability level per source. Readability statistics were calculated using the Flesch-Kincaid Grade Level Readability formula. Readability is calculated using a formula where the total number of syllables, words, and sentences in a given passage are determinants of its difficulty. The obtained value is intended to match up with the grade level at which one should be able to comfortably read the passage [@Kincaid1975]. For example, a text with a readability score of 11 should be easily read by a high school junior.

As seen in Table 1, the sources are similar in some aspects yet different in others. Valence appears to be slightly positive across all sources. The standard deviations seem to indicate little to no presence of a difference in valence. 
*The New York Times* published the greatest number of articles as well as total words. *Breitbart* featured the lowest number of articles, and *NPR* the lowest number of total words from all articles. Per indvidual article, however, *Breitbart* appears to feature the highest average number of words as well as unique words. Once again the standard deviations call into question any apparent differences between sources. Finally, *Fox News* appears to be the most readable source on average, while *The New York Times* is the least readable. This might be attributable to the greater number of tokens in the average *New York Times* article compared to *Fox News*. As before, the standard deviations for readability hamper one's ability to draw solid conclusions regarding differences in the sources' use of language.

@@

The multilevel model did not indicate the presence of any significant or practical effect of political lean for any of the five moral foundations. The strongest effect size was observed for the *authority/respect* foundation, but the effect was in the opposite direction from what was originally hypothesized - liberal sources tended to use more *authority/respect* words than did conservative sources. Descriptive and test statistics, *p*-values and effect sizes (Cohen's *d*) can be found in Table 2. To interpret the weighted scores, one can examine the mean and standard deviations for each. A zero score for the mean, with a non-zero standard deviation, would indicate a perfect balance of positive and negative words in each category, likely representing a neutral tone when all words are considered. Negative percentages would indicate more representation of the negative words in the MFD area, while positive percentages indicate an endorsement of the positive words in a MFD. Therefore, we suggest using the sign of the mean score to determine the directionality of the endorsement for the MFD (positive, neutral, negative), and the standard deviation to ensure that a zero score is not zero endorsement (i.e., a SD of zero indicates no words were used). Based on the weighted percent values for the five foundations, the researchers observed that MFD words seem to make up a small portion of the article text. Furthermore, the observed percentages and means appear to indicate a generally positive endorsement of all five foundations across both liberal and conservative sources.

@@

#Kav Table
####build the table here####
tableprint = matrix(NA, nrow=10, ncol=11)
colnames(tableprint) = c("Source", "Mean_Valence", "SD_Valence", "Number_of_Articles", "Total_Words", "Mean_Tokens", "SD_Tokens", "Mean_Types", "SD_Types", "Mean_Readability", "SD_Readability")

tableprint[ , 1] = print(unique(final$Source))
tableprint[ , 2] = tapply(final_kav$valenceavg, final_kav$Source, mean)
tableprint[ , 3] = tapply(final_kav$valenceavg, final_kav$Source, sd)
tableprint[ , 4] = tapply(final_kav$X, final$Source, length)
tableprint[ , 5] = tapply(final_kav$wordcount, final_kav$Source, sum)
tableprint[ , 6] = tapply(final_kav$wordcount, final_kav$Source, mean)
tableprint[ , 7] = tapply(final_kav$wordcount, final_kav$Source, sd)
tableprint[ , 8] = tapply(final_kav$uniquewords, final_kav$Source, mean)
tableprint[ , 9] = tapply(final_kav$uniquewords, final_kav$Source, sd)
tableprint[ , 10] = tapply(final_kav$FK, final_kav$Source, mean)
tableprint[ , 11] = tapply(final_kav$FK, final_kav$Source, sd)

#change this to match
apa_table(as.data.frame(tableprint),
          note = "Readability statistics were calculated using the Flesch-Kincaid Grade Level readability formula.",
          caption = "Kavanaugh - Descriptive Statistics by Source", 
          col.names = c("Source", "$M_Valence", "$SD_Valence$", "Number of Articles", "Total Words", "$M_Tokens$", "$SD_Tokens$", "$M_Types$", "SD_Types", "$M_Readability$", "$SD_Readability$"),
          format = "latex",
          escape = F,
          align = c("l", rep("c", 6))
          )

@@

tapply(final_gs$FK, final_gs$Source, mean)
tapply(final_gs$FK, final_gs$Source, sd)

#GS Table
####build the table here####
tableprint = matrix(NA, nrow=10, ncol=11)
colnames(tableprint) = c("Source", "Mean_Valence", "SD_Valence", "Number_of_Articles", "Total_Words", "Mean_Tokens", "SD_Tokens", "Mean_Types", "SD_Types", "Mean_Readability", "SD_Readability")

tableprint[ , 1] = print(unique(final$Source))
tableprint[ , 2] = tapply(final_gs$valenceavg, final_gs$Source, mean)
tableprint[ , 3] = tapply(final_gs$valenceavg, final_gs$Source, sd)
tableprint[ , 4] = tapply(final_gs$X, final$Source, length)
tableprint[ , 5] = tapply(final_gs$wordcount, final_gs$Source, sum)
tableprint[ , 6] = tapply(final_gs$wordcount, final_gs$Source, mean)
tableprint[ , 7] = tapply(final_gs$wordcount, final_gs$Source, sd)
tableprint[ , 8] = tapply(final_gs$uniquewords, final_gs$Source, mean)
tableprint[ , 9] = tapply(final_gs$uniquewords, final_gs$Source, sd)
tableprint[ , 10] = tapply(final_gs$FK, final_gs$Source, mean)
tableprint[ , 11] = tapply(final_gs$FK, final_gs$Source, sd)

#change this to match
apa_table(as.data.frame(tableprint),
          note = "Readability statistics were calculated using the Flesch-Kincaid Grade Level readability formula.",
          caption = "Government Shutdown - Descriptive Statistics by Source", 
          col.names = c("Source", "$M_Valence", "$SD_Valence$", "Number of Articles", "Total Words", "$M_Tokens$", "$SD_Tokens$", "$M_Types$", "SD_Types", "$M_Readability$", "$SD_Readability$"),
          format = "latex",
          escape = F,
          align = c("l", rep("c", 6))
          )
```
